---
title: "Concomitant vaccination of Meningococcal B with infant immunisation schedules"
subtitle: "Analysis of Medically Attended events"
date: "`r Sys.Date()`"
output: 
 bookdown::pdf_document2:
   toc: no
   fig_align: center
   fig_height: 4
   fig_width: 6
---

\tableofcontents
\addcontentsline{toc}{section}{\listfigurename}

\clearpage

\listoffigures

\clearpage

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
 	message = FALSE,
  warning = FALSE,
 	fig.pos = "!htbp",
  out.extra = ""
)

## read in the libraries

library(cmdstanr)
library(posterior)
library(ggplot2)
library(tidyverse)
library(stringr)
library(data.table)
library(here)

## set the seed

set.seed(12856421)

## colourblind friendly palette

colours <- c("#88CCEE", "#CC6677", "#DDCC77")

## load data



```

# Outline

This document details the analysis of medically attended events for the work investigating the reactogenicity of the Meningococcal B (MenB) vaccine given either concomitantly or separately to infant immunisation schedules on the National Immunisation Program (NIP). The following sections detail the notation and modelling approaches. To simplify the examples we assume there are no covariate effects on the outcome and that there is only one scheduled timepoint.

# Notation

Suppose that there are $N$ (eligible) responders to the AusVaxSafety survey soliciting reactogenicity information within three days of vaccination for infants who received a MenB vaccine and their NIP scheduled vaccines either concomitantly or separately. We denote these individual responders with $i \in \{1,2,\ldots,N\}$.

We denote the vaccination *strategy* for responder $i$ with $s_i \in \{\text{Concomitant} = 1, \text{Separate} = 2\}$ representing whether the vaccinations were delivered concomitantly or separately. Responder $i$ has one survey response if $s_i = \text{Concomitant}$ and two survey responses if $s_i = \text{Separate}$. The scheduled NIP timepoints for responder $i$ are denoted $t_i \in \{\text{2 months} = 1, \text{4 months} = 2, \text{6 months} = 3, \text{12 months} = 4\}$. Covariates sex, Indigenous status and comorbidity for responder $i$ are denoted $w_i \in \{\text{Male} = 0, \text{Female} = 1\}$, $x_i \in \{\text{Non-Indigenous} = 0, \text{Indigenous} = 1\}$ and $z_i \in \{\text{No} = 0, \text{Yes} = 1\}$, respectively.

The event of interest (medically attended adverse events) is defined in the following section.

# Modelling

## Scenario A

In Scenario A, the objective of the model is to estimate the probability of **at least one** medically attended event reported by responders receiving each strategy. As an event may only occur once per responder, i.e., either no events or at least one event, we denote MA for responder $i$ with $y_i \in \{\text{Zero MA} = 0, \text{At least one MA} = 1\}$. The model is:

\begin{align*}
y_i &\sim \text{Bernoulli}(p_i) \\
\text{logit}(p_i) &= \mu_{s_i,t_i} + w_i \times \beta + x_i \times \gamma + z_i \times \delta \\
\mu_{s_i,t_i} &\sim \text{N}(a,b) \\ 
\beta, \gamma, \delta &\sim \text{N}(0,1)
\end{align*}

Here, $p_i$ is the probability of responder $i$ reporting at least one MA and $\mu_{s_i,t_i}$ is the log odds of a responder receiving strategy $s_i$ at scheduled timepoint $t_i$ reporting at least one event. The log odds ratios for sex, Indigenous status and comorbidity are represented by $\beta$, $\gamma$ and $\delta$, respectively. The hyperparameters $a$ and $b$ would be chosen so that the prior distributions are weakly informative and will depend on the specific outcome. For each scheduled timepoint $t_i$, the difference in event probabilities can be calculated with $\text{logit}^{-1}(\mu_{\text{Concomitant},t_i}) - \text{logit}^{-1}(\mu_{\text{Separate},t_i})$.


\small
```{r, echo = FALSE, comment = ""}

#dat <- read.csv()
dat <- read.csv(file = "C:/Users/ETay/Documents/Work documents/AVS work/Thuy_concom/dat_modelA.csv")

N_A  = nrow(dat)                     ## number of responders
N_strat_A  = 2                      ## number of strategies #3
N_sched_A  = 4                       ## number of schedules

#s_A <- dat$vax_sequence              ## vaccine strategy - "Concomitant vaccination" = "1", "NIP first" = "2", "MenB first" = "3"
s_A <- dat$group                     ## vaccine strategy - "Concomitant vaccination" = "1", "Separate" = "2"
t_A <- dat$schedule                  ## schedule - 1 = 2 months, 2 = 4 months, 3 = 6 months, 4 = 12 months
w_A <- dat$sex                       ## sex - 0 = "Male", 1 = "Female"
x_A <- dat$indig                     ## Indigenous status -0 = Non-indig, 1 = Aboriginal and Torres Strait Islander
z_A <- dat$pmh                       ## comorbidity - 0 = No, 1 = Yes
y_A <- dat$medical_attention         ## outcome - MA - 0 = No, 1 = Yes


a <-  qlogis(0.3)           ## prior distribution mean - depends on the schedule and the vaccine strategy?
b <-   2                    ## prior distribution standard deviation

dat_A <- list(N = N_A,
              N_strat = N_strat_A,
              N_sched = N_sched_A,
              s = s_A,
              t = t_A,
              w = w_A,
              x = x_A,
              z = z_A,
              y = y_A,
              a = a,
              b = b)

#writeLines(readLines("C:/Users/mdymock/Documents/Projects/AVS/Concomitant/model_A.stan"))
#writeLines(readLines("C:/Users/etay/Documents/Work documents/AVS work/Thuy_concom/model_A.stan"))
```
\normalsize

\clearpage

Fit the model, extract the posterior distributions and visualise:

```{r, echo = FALSE}
#model_A <- cmdstan_model("C:/Users/mdymock/Documents/Projects/AVS/Concomitant/model_A.stan")
model_A <- cmdstan_model("C:/Users/etay/Documents/Work documents/AVS work/Thuy_concom/AVS_Concom_pub/model_A.stan")
drp <- capture.output({fit_A <- model_A$sample(dat_A, chains = 8, parallel_chains = 8)})
draws_A <- as_draws_matrix(fit_A$draws(c("mu")))
dat_vis_A <- data.frame(x = plogis(as.vector(draws_A)),
                        Parameter = rep(c("Concomitant", "Separate"), each = 8000))
```

```{r, echo = FALSE, fig.cap = "Posterior distributions for Scenario A parameters."}
ggplot(dat_vis_A, aes(x = x, colour = Parameter)) +
  geom_density() +
  scale_colour_manual(values = colours)
```

The means and 95\% credible intervals for the event probabilities for each strategy are: 

```{r, echo = FALSE}
summary_fun_A <- function(strat, dat){
  mn <- format(round(mean(dat[dat$Parameter == strat,]$x), 2), nsmall = 2)
  cr <- format(round(quantile(dat[dat$Parameter == strat,]$x, c(0.025, 0.975)), 2), nsmall = 2)
  paste0(strat, " strategy: ", mn, " (", cr[1], ", ", cr[2], ")")
}

summary_fun_A("Concomitant", dat_vis_A)
summary_fun_A("Separate", dat_vis_A)
```

\clearpage

